---
title: "REI_test"
author: "Maya Green"
date: "2026-03-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Important note: all of the tiff files (rasters) generated by the wind interpolation function were blank for some reason, so all of this below is creating a bunch of blank layers. Need to go back to the drawing board to figure out:
1. Why the tif files were generated with all NAs
2. Why the resulting data from Copernicus is so low resolution (16 km)
  answer: it is 0.25 deg resolution, which is standard for this data. This is why we interpolate.  


```{r}
#######################################################################
##########Environmental Data Layers for SG sdm#########################
#######################################################################

#### Load Packages ####
devtools::install_github('jgcri/rgis')

library(dplyr)
library(purrr)
library(tidyr)
library(sf)
library(terra)
library(rgis)
library(raster)
install.packages("prioritizr")
library(prioritizr)

# Housekeeping ----

# Speed up raster processes

rasterOptions(chunksize = 1e+05, maxmemory = 1e+09)

# Read in interpolated wind data ----

# Raster stack of wind frequency by direction
 
wind_freq <- list.files(path = "processed_evd/processed_evd/",
                   pattern = "^freq.+tif$",
                   full.names = T) %>% 
  stack(.) 

# Raster stack of average daily maximum wind speed by direction

wind_spd <- list.files(path = "processed_evd/processed_evd/",
                  pattern = "^mx.+tif$",
                  full.names = T) %>% 
  stack(.) 

# Combine fetch data and wind data ----
# make combined fetch shp into a .rds
combined_fetch <- st_read("processed_evd/fetch/combined_fetch_eff.shp")
saveRDS(combined_fetch, file = "processed_evd/fetch/combined_fetch_eff.rds")


# Read in effective fetch data

Fetch_eff <- readRDS("processed_evd/fetch/combined_fetch_eff.rds")
```

```{r, checking extent and crs overlaps}
# put everything into terra format
wind_freq <- terra::rast(wind_freq)
wind_spd <- terra::rast(wind_spd)

# project the rasters to the same crs as fetch_eff data 
wind_freq <- project(wind_freq, crs(Fetch_eff))
wind_spd <- project(wind_spd, crs(Fetch_eff))

# crop extents 
wind_freq <- crop(wind_freq, extent(Fetch_eff))
wind_spd <- crop(wind_spd, extent(Fetch_eff))


# Extract frequency and speed values for fetch locations

extract_freq <- fast_extract(wind_freq, Fetch_eff) 
extract_spd <-fast_extract(wind_spd, Fetch_eff)

wind_combine <- left_join(extract_freq, extract_spd, by = "id") # this is not working 

```

```{r}
# I am suspicious
# it appears that the .tif rasters are all blank??
grey <- raster("processed_evd/processed_evd/freq_180_spline.tif")
plot(grey)
mx <- raster("processed_evd/processed_evd/mx_spd_135_spline.tif")
plot(mx)


# this worked, the previous did not
bathy <- raster("environmental_dat/Chissassibi_SDB_Sentinel2.tif")
plot(bathy)

```



```{r}
# Calculate Relative Exposure Index ----

expo_by_site <- wind_combine %>%
  rowwise() %>% 
  mutate(freq_total = list(cbind(c_across(contains("freq")))),
         max_wind = list(cbind(c_across(contains("mx")))),
         wind_dir_freq = list(tibble(max_wind, 
                                freq_total, 
                                direction = c(135,180,225,270,315,360,45,90)))) %>%
  ungroup() %>% 
  mutate(., wind_dir_freq = map(wind_dir_freq, function(.data) {
    .data %>% 
      dplyr::arrange(., direction)
  }),
  fetch_eff = map(fetch_eff, function(.data){
    .data %>% 
      dplyr::arrange(., direction)
  })) %>% 
  mutate(REI = map2_dbl(wind_dir_freq, fetch_eff, # calculate relative exposure
                        ~sum((.x$max_wind*.x$freq_total*.y$fetch), na.rm = T))) %>%  
  dplyr::select(site_names, REI) # keep site, REI, and geometry

# Save REI object

saveRDS(expo_by_site, 'Data/REI.rds')

# Write to csv

st_write(obj = expo_by_site,
         dsn = 'Data/REI.csv',
         layer_options = 'GEOMETRY=AS_XY')

```

