---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---

```{r}
install.packages("tidyverse")
```


```{r}
# load packages

library(dplyr)
library(tidyverse)
library(readr)
library(tidyr)
library(sf)
library(raster)
#library(INLA)
library(terra)
library(ggmap)
library(stringi)
library(purrr)
library(units)
library(parallel)
library(multidplyr)

```
# First download canada coastlines data from https://open.canada.ca/data/en/dataset/6c78fb2f-d23b-45b4-b3af-cc6f6cc4fff8

```{r}
# make a polygon of the James Bay coastline including islands

canada_coast <- read_sf(dsn = "./canada_coastline_polygons/lhy_000h16a_e.shp")
canada_coast # data is currently in Lambert Conformal Conic form
```

Processing this data to run it using a fetch function created by O'Brien. Below are the steps needed to create each input needed for the function.

# Creating the necessary variables for fetch calculation 
# polygon_layer

```{r}
#'polygon_layer = sf object where the
#'                polygon geometries represent any obstructions to fetch
#'                calculations including the coastline, islands and/or 
#'                exposed reefs

polygon_layer <- read_sf("./processed_EvD/shapefiles/james_bay_poly_coastline.shp")
polygon_layer <- st_as_sf(polygon_layer)

polygon_layer # projected CRS is NAD83/Statistics Canada Lambert
```


# site_layer
I created a ton of point sites in a shapefile earlier so we will use those for site layer, but we need to narrow these down (there are >400,000 points in this layer)

```{r}
#' site_layer = sf object where the point 
#'                geometries represent the site locations.

site_layer <- read_sf("./") # this shp has about 450,000 points at a 300m resolution along the entire coast

#' so that's 450,000*32 or 14,400,000 raw fetch vectors 
#' which ends up giving 3.6 million effective fetch vectors 
#' that's... a lot


site_layer <- st_as_sf(site_layer)
site_layer <- site_layer %>% 
  dplyr::select(geometry)

# project to polygon CRS
site_layer <- st_transform(site_layer, st_crs(polygon_layer))
site_layer


```

# max_dist, n_directions, and site_names
```{r}
#' max_dist = numeric. Maximum distance in kilometres (default 300). This 
#'                will need to be scaled manually if the units for the CRS are 
#'                not 'm'.
# 300 m makes sense for James Bay

#' n_directions = numeric. The number of fetch vectors to calculate per 
#'                quadrant (default 8). 
#'                
#' site_names = character vector of the site names. If missing, the site 
#'                names are taken from a column of the data associated with 
#'                site_layer matching the regular expression 
#'                {^[Nn]ames{0,1}}. If there is no such column, then 
#'                default names are created ('Site 1', 'Site 2', ...).

site_names <- c(1:20) # length of site_layer
site_names <- as.character(site_names)

#' quiet = logical. Suppress diagnostic messages? (Default {FALSE}).


#one final check to make sure every variable looks good (site and polygon layers should have same CRS)
polygon_layer
site_layer
site_names

# need to also make site_vars for the parallel function
site_layer$site_var <- NA
site_var <- site_layer$site_var
```

# Look at this section Joe 

This is one section that will take a lot of compute power.
This function calculates fetch in 36 directions, so it's 32*number of points I give it (site_layer).
Afterwards we calculate effective fetch which averages vectors across the 8 cardinal directions (NSEW etc).
So total number of *effective* fetch points is fetch_raw_points/4

(see parallel processing fetch function at line 268)

```{r}
fetch = function(polygon_layer, site_layer, max_dist = 300, n_directions = 8,
                 site_names, quiet = FALSE){
  
  if (!is(polygon_layer, "sf") && !is(st_geometry(polygon_layer), c('sfc_POLYGON', 'sfc_MULTIPOLYGON')))
    stop(paste("polygon_layer must be a sf object with a POLYGON or MULTIPOLYGON geometry type.\nSee",
               "'?sf' for details on how to create a",
               "sf object."), call. = FALSE)
  
  if (!is(site_layer, "sf") && !is(st_geometry(site_layer), c('sfc_POINT', 'sfc_MULTIPOINT')))
    stop(paste("polygon_layer must be a sf object with a POINT or MULTIPOINT geometry type.\nSee",
               "'?sf' for details on how to create a",
               "sf object."), call. = FALSE)
  
  if(is(st_geometry(polygon_layer), 'sfc_MULTIPOLYGON'))
     st_cast(polygon_layer, 'POLYGON')
  
  if (!is.numeric(max_dist) || length(max_dist) != 1)
    stop("max_dist must be a single number.", call. = FALSE)
  
  if (!is.numeric(n_directions) || length(n_directions) != 1)
    stop("n_directions must be a single integer.", call. = FALSE)
  n_directions = round(n_directions)
  
  if (n_directions < 1 || n_directions > 90)
    stop("n_directions must be between 1 and 90.", call. = FALSE)
  
  if (!missing(site_names)){
    site_names = as.character(site_names)
    
    if (length(site_names) != nrow(site_layer)){
      warning(paste("lengths differ for the number of sites and site names;", 
                    "using default names instead."), call. = FALSE)
      site_names = paste("Site", seq_along(st_geometry(site_layer)))
    }
  } else {
    if (any(grepl("^[Nn]ames{0,1}$", names(site_layer)))){
      name_col = grep("^[Nn]ames{0,1}$", names(site_layer))[[1]]
      site_names = as.character(pull(site_layer, name_col))
    } else {
      site_names = paste("Site", seq_along(st_geometry(site_layer)))
    }
  }
  
  quiet = as.logical(quiet[1])
  
  ## Check if the polygon and points layers are projected, and ensure they have 
  ## the same CRS.
  which_proj = list(polygon_layer, site_layer) %>% 
    map(., st_crs) %>% 
    map_lgl(., ~grepl('PROJCRS',.$wkt[1]))
  
  if (all(!which_proj))
    stop("polygon_layer and/or site_layer must be projected to calculate fetch", 
         call. = FALSE)
  
  if (all(which_proj) && !identical(st_crs(polygon_layer), st_crs(site_layer))){
    warning("the CRS for polygon_layer and site_layer differ; transforming
              site_layer CRS to match")
    site_layer = st_transform(site_layer, st_crs(polygon_layer))
  }
  
  if (!which_proj[1]){
    if (!quiet)
      message("projecting polygon_layer onto the site_layer CRS")
    polygon_layer = st_transform(polygon_layer, st_crs(site_layer))
  }
  
  if (!which_proj[2]){
    if (!quiet)
      message("projecting site_layer onto the polygon_layer CRS")
    site_layer = st_transform(site_layer, st_crs(polygon_layer))
  }
  
  if (!quiet)
    message("checking site locations are not on land")
  
  # Should remove these sites with warning instead of returning error
  if (any(lengths(st_intersects(site_layer, polygon_layer)) > 0))
    stop("at least one site location is on land")
  
  # Convert max_dist to appropriate units.
  # First of all convert max_dist to metres (default)
  max_dist = max_dist * 1000
  
  # Double check if metres are the correct units
  # If not, warn the user that the supplied max_dist should be scaled 
  # appropriately
  if (st_crs(polygon_layer)$units != "m")
    warning("the PROJ.4 unit is not metres; ensure max_dist has been scaled 
            appropriately")
  
  directions = head(seq(0, 360, by = 360 / (n_directions * 4)), -1)
  
  # Rearrange sequence order to start at 90 degrees to match up with the output
  # from st_buffer
  directions = unlist(split(directions, directions < 90), use.names = FALSE)
  
  # Create nested data that calculates and stores Fetch Objects and lengths for
  # each site
  
  message("calculating fetch for ", nrow(site_layer), " sites")
  
  fetch_sf <- site_layer %>%
    bind_cols(., tibble(site_names)) %>%
           # Create polygon (approximating a circle) with a given radius. These 
           # vertices are used for creating the end points for the fetch vectors.
    mutate(d_bff = st_buffer(geometry, dist = max_dist, nQuadSegs = n_directions),
           # Calculate end points at the maximum distances.       
           fetch_ends = map(d_bff, ~head(st_coordinates(.)[,1:2], -1)),
           fetch_ends = map(fetch_ends, ~.[order(directions),]),
           # Create matrix repeating site coordinates (nrow = number of fetch directions)
           coords = map(geometry, st_coordinates),
           coords = map(coords, ~.[rep(1:nrow(.), times = n_directions * 4),]),
           # bind site coordinates with fetch ends
           all_mat = map2(coords, fetch_ends, rbind),
           # split matrix into list
           all_list = map(all_mat, ~split(., sort(directions))),
           # Create a list of 2x2 matrices indicating the start and end coordinates
           line_list = map(all_list, ~map(., matrix, ncol = 2)),
           # create sf line objects for all fetch directions
           # Lines radiate around site to max distance
           fetch_st_lines = map(line_list, ~st_sf(geom = st_sfc(st_multilinestring(.), 
                                                   crs = st_crs(polygon_layer)))),
           # determine indices of subset of polygons intersecting 
           # with full length fetch lines
           poly_sub = map(fetch_st_lines, ~st_intersects(., polygon_layer)),
           # cast from multiline to line (one row per direction)
           fetch_st_lines = map(fetch_st_lines, st_cast, 'LINESTRING'),
           # sort rows in sf line dataframes by fetch direction
           fetch_st_lines = map(fetch_st_lines, ~ mutate(., direction = sort(directions))),
           # Erase polygon layer from fetch lines
           fetch_st_lines = map2(fetch_st_lines, poly_sub,
                                ~st_difference(.x, st_union(st_combine(polygon_layer[unique(unlist(.y)),])))),
           # Extract first line segment from each fetch direction (this is fetch line shortened by first polygon intersection)
           fetch_st_lines = map(fetch_st_lines, ~st_cast(., 'LINESTRING')[as.character(seq(1:(n_directions * 4))),]),
           # Calculate fetch line lengths
           fetch_length = map(fetch_st_lines, st_length)) %>%
    dplyr::select(., site_names, fetch_st_lines, fetch_length) %>%
    unnest(., cols = c(fetch_length, fetch_st_lines)) %>% 
    dplyr::select(-geom)

}

```

# parallel processing fetch function

```{r}
############################################################
####    Run fetch function with parallel processing     ####
############################################################

# source fetch function

# source("./processed_EvD/fetch/obrienjm25-REI-WaveExp-fbf17f9/Code/3Fetch.R")

# Run fetch function in parallel using multidplyr

#' polygon_layer = sf object where the
#'                polygon geometries represent any obstructions to fetch
#'                calculations including the coastline, islands and/or 
#'                exposed reefs.
#' site_layer = sf object where the point 
#'                geometries represent the site locations.
#' max_dist = numeric. Maximum distance in kilometres (default 300). This 
#'                will need to be scaled manually if the units for the CRS are 
#'                not 'm'.
#' n_directions = numeric. The number of fetch vectors to calculate per 
#'                quadrant (default 8).
#' site_var (character) = column name in site_layer sf object holding variable
#'                        for site names

fetch_parallel = function(polygon_layer, site_layer, max_dist = 300, 
                          n_directions = 8, site_var){ 
  
  # Determine number of cores available minus 1
  cl <- detectCores() - 1
  
  # Set up new cluster for parallel processing with one fewer than number of
  # available cores
  # Load required libraries, objects, and functions to each core
  cluster <- new_cluster(n = cl) %>% 
    cluster_library(c('sf','raster','stringi','purrr','dplyr','tidyr','units')) %>% 
    cluster_assign(fetch = fetch,
                   site_layer = site_layer,
                   polygon_layer = polygon_layer,
                   max_dist = max_dist,
                   n_directions = n_directions,
                   site_var = site_var)
  
  # Calculate fetch for all survey sites in 11.25 degree increments
  fetch_20points_c <- site_layer %>%
    bind_cols(., cl_group = rep(1:cl, length.out = nrow(.))) %>% # grouping variable to split obs over cores
    group_by(cl_group) %>% # group by grouping variable
    partition(cluster = cluster) %>% # partition obs. to clusters (create party_df)
    do(fetch(polygon_layer = polygon_layer, # calculate fetch in parallel
             site_layer = .,
             max_dist = max_dist,
             n_directions = n_directions,
             site_names = pull(., var = site_var))) %>% 
    collect() %>% # combine observations from all clusters
    st_as_sf(sf_column_name = 'geometry') %>% 
    ungroup()
  
  return(fetch_20points_c)
  
}

```

Calculating effective fetch from the csv generated above
Feel free to ignore most of the commented-out code, I have adapted it to work for my data structure

```{r}
#######################################################################
########## Calculate effective fetch from raw fetch vectors ###########
#######################################################################

# Description: ----

# This script takes fetch distances in 32 directions around grid points,
# the output of Python fetch script, and calculates for each point:
#
# 1) Minimum fetch: the shortest fetch distance (proxy for distance to coast)
# 2) Sum fetch: the sum of all 32 fetch distances (proxy for wave exposure)
# 3) Effective fetch: in 8 directions (N, NE, E, SE, S, SW, W, NW); 
#    needed to calculate Relative Exposure Index (REI)*
#
# *For description of effective fetch and REI, see:
#
# Fonseca MS, Bell SS (1998) Influence of physical setting on seagrass
#   landscapes near Beaufort, North Carolina, USA. Mar Ecol Prog Ser
#   171:109-121

# Requirements: ----

# csv file(s) containing outputs of Python fetch script
# i.e. fetch distance (metres) in 32 directions around grid points
# One row per point

# Load packages ----

library(dplyr)
library(purrr)
library(tidyr)
library(data.table)
library(sf)
library(vroom)

# Source functions ----

# source('./processed_EvD/fetch/obrienjm25-REI-WaveExp-fbf17f9/Code/roll_recycle_fun.R')

# Controls ----

# relative contribution of fetch vectors surrounding each heading to calculation of effective fetch

weights <- cos(c(45,33.75,22.5,11.25,0,11.25,22.5,33.75,45) * (pi/180))

# Path to project directory (string)

projDir <- "~/Desktop/green-bces"

# Coordinate reference system of fetch calculations (integer EPSG code)

crs_fetch <- 3347 # NAD83 / Statistics Canada Lambert

# Combine fetch data from all points into one data table ----

# list of fetch files

#fetch_csv <- list.files(path = projDir,
                        #pattern = 'csv',
                        #recursive = TRUE,
                        #full.names = TRUE)

# Combine csv files from separate clusters of points and reshape to long format

#fetch_summary <- map(fetch_csv, fread) %>%
  #rbindlist(idcol = 'cluster') %>%
 # mutate(., site_names = seq_along(cluster)) %>%
  #pivot_longer(data = .,
               #cols = starts_with("bearing"),
               #values_to = "fetch_length_m",
               #names_to = "direction",
               #names_prefix = "bearing",
               #names_transform = list(direction = as.numeric)) %>%
  #setDT()


# site names column was full of NAs 
site_names_vec <- rep(c(1:20), each = 36)
# raw_fetch_36dir$site_names <- site_names_vec
# raw_fetch_36dir$site_names <- factor(raw_fetch_36dir$site_names)
```

This is from a previous attempt at doing a fetch subsample, fetch_20points no longer exists

```{r}
# fetch_proxies <- fetch_20points %>% 
  #group_by(geometry) %>% 
  #mutate(min_fetch = min(fetch_length), sum_fetch = sum(fetch_length))

# Calculate minimum fetch (proxy for distance to coast) and sum fetch (proxy for wave exposure)
#fetch_proxies <- fetch_20points[, .(min_fetch = min(fetch_length),
                                # sum_fetch = sum(fetch_length)),
                             #by = .(site_names,"geometry")]

#vroom_write(fetch_proxies, "Data/fetch_proxies.csv", delim = ",")

# Calculate effective fetch ----

#fetch_eff <- fetch_20points %>%
  #select(geometry, direction, fetch_length) %>% 
  #group_by(geometry) %>% 
  #nest() %>% 
  #bind_cols(., site_names = as.character(seq_len(nrow(.)))) %>%   
  # Calculate effective fetch and place in tibble
  #mutate(fetch_eff = map(data, ~roll.recycle(.$fetch_length, 9, 8, by = 4)),
         #fetch_eff = map(fetch_eff, ~as.vector((weights %*% .)/sum(weights))),
         #fetch_eff = map(fetch_eff, ~tibble(direction = c(360, seq(45, 315, 45)), fetch = .))) %>% 
  #st_set_geometry("geometry")

#saveRDS(fetch_eff, "Data/fetch_effective.rds")
```

Same note applies here

```{r}
# calculating effective fetch

#fetch_eff <- fetch_20points %>%
 # group_by(site_names) %>%
 # summarise(
    #fetch_eff = list({
    #  f <- roll.recycle(fetch_length, 9, 8, by = 4)
    #  f <- as.vector((weights %*% f) / sum(weights))
    #  tibble(
     #   direction = c(360, seq(45, 315, 45)),
    #    fetch = f
  #    )
 #   }),
 #   .groups = "drop"
 # ) %>%
 # tidyr::unnest(fetch_eff)

# visualize fetch using a raster
# fetch_raster <- rast(fetch_eff)
# plot(fetch_eff)

```

Relative Exposure Index calculation

```{r}
install.packages("weathercan", repos = "https://dev.ropensci.org")
library(weathercan)
install.packages("naniar")
library(naniar)

# loading in wind data to get started on the REI calculation

# fiding stations on the east coast of JB
east_coast <- stations_search(coords = c(53.547211, -78.984518),
                              interval = "day", dist = 100)

# getting weather data from these for 2025
east_coast_data <- weather_dl(station_ids = east_coast$station_id, 
                              start = "2025-12-01", 
                              interval = "hour", quiet = TRUE) 

# only shows two stations because they are the only ones with wind data going up until 2025

# finding stations on the west coast of JB 
west_coast <- stations_search(coords = c(53.474720, -82.139678),
                              interval = "day", dist = 100)

# getting west coast weather data
west_coast_data <- weather_dl(station_ids = west_coast$station_id, 
                              start = "2025-12-01", 
                              interval = "hour", quiet = TRUE)
# same with this wind station

east_coast_data <- as.data.frame(east_coast_data)

east_coast_data_wind <- east_coast_data %>% 
  dplyr::select(station_name, station_id, lat, lon, date, time, wind_dir, wind_spd) %>% 
  filter(!is.na(wind_dir & wind_spd)) %>% 
  dplyr::group_by(date) %>% 
  mutate(mean_dir = mean(wind_dir)) %>% 
  mutate(wind_dir_360 = wind_dir*10) 
  
# calculate wind frequency
wind_freq_data <- east_coast_data_wind %>% 
  group_by(date, station_name, wind_dir) %>% 
  summarise(n = n()) %>% 
  mutate(wind_freq = n/24*100)

west_coast_wind_data <- west_coast_data %>% 
  dplyr::select(station_name, station_id, lat, lon, date, time, wind_dir, wind_spd) %>%
  filter(!is.na(wind_dir & wind_spd)) %>% 
  dplyr::group_by(date) %>% 
  mutate(mean_dir = mean(wind_dir)) %>% 
  mutate(wind_dir_360 = wind_dir*10)
  

hist(east_coast_data_wind$mean_dir)
hist(west_coast_wind_data$mean_dir)


```


